Interactive Visualization of Hierarchically Structured Data

### 2017-03-18 treelapse.Rmd

In the tree structured bacterial dynamics problem, the goal is to
describe changes in bacterial abundances in an environment over
time. As in the differential abundance problem, it is useful if these
descriptions can be given at the highest subtree at which the pattern
appears. Specific questions of interest often have an ecological
flavor. For example, researchers are often interested in understanding
how bacterial populations respond to sudden or gradual environmental
changes or how species fill, drop out from, or compete for
environmental niches. Medically, these questions are important for
illuminating the impact of antibiotic time courses or diet changes,
for example.

@@

```{r antibioticoverview, fig.cap = "The two antibiotic time courses are readily apparent even when only highlighting the most abundant taxa."}
include_graphics("figure/antibiotic_overview.png")
```

@@

```{r zillowsf, fig.cap = "To study teh range in home prices within San Francisco County, we can select the associated nodes using a treebox."}
include_graphics("figure/zillow_sf.png")
```

@@

```{r gptimebox, out.width = "200px", fig.show = "hold", fig.cap = "An application to the Global Patterns demonstrates how linking in treelapse can be applied to combine hierarchical clustering and parallel coordinates views."}
include_graphics(c("figure/gp_cluster1.png", "figure/gp_cluster2.png"))
```

@@

In summary, we have built an easily accessible R package for
visualization techniques in a very specific methodology problem –
analysis of differential abundance and dynamics in hierarchically
structured data – that appears in a variety of application domains. We
have leveraged a link between R and D3 [@vaidyanathan2014htmlwidgets]
to create exploratory visualizations while still deep in data
analysis; in this way our work is a departure from the culture of
polished, journalistic visualizations prioritized by the D3
community. Finally, we have given a series of examples to demonstrate
how the general visualization techniques of focus-plus-context and
linked brushing can be practically incorporated into a range of
practical analysis workflows, from studying the impact of bacteria on
human health to better allocating units in commuter bikesharing
systems.

>>>>

In the tree-structured bacterial dynamics problem, the goal is to
describe changes in bacterial abundances in an environment over
time. As in the differential abundance problem, it is useful if these
descriptions can be given at the highest subtree at which the pattern
appears. Specific questions of interest often have an ecological
flavor. For example, researchers are often interested in understanding
how bacterial populations respond to sudden or gradual environmental
changes or how species fill, drop out from, or compete for
environmental niches. Medically, these questions are important for
illuminating the impact of antibiotic time courses or diet changes,
for example.

@@

```{r antibioticoverview, fig.cap = "The two antibiotic time courses are readily apparent even when only highlighting the most abundant taxa. We have annotated the figure with its three main components."}
include_graphics("figure/annotated_antibiotic_overview.png")
```

@@

```{r zillowsf, fig.cap = "To study the range in home prices within San Francisco County, we can select the associated nodes using a treebox."}
include_graphics("figure/zillow_sf.png")
```

@@

```{r gptimebox, out.width = "250px", fig.show = "hold", fig.cap = "An application to the Global Patterns demonstrates how linking in treelapse can be applied to combine hierarchical clustering and parallel coordinates views."}
include_graphics(c("figure/gp_cluster1.png", "figure/gp_cluster2.png"))
```

@@

In summary, we have built an easily accessible R package for
visualization techniques in a very specific methodology problem –
analysis of differential abundance and dynamics in hierarchically
structured data – that appears in a variety of application domains. We
have leveraged a link between R and D3 [@vaidyanathan2014htmlwidgets]
to create visualizations during the exploratory phase of data analysis; in this
way our work is a departure from the culture of polished, journalistic
visualizations prioritized by the D3 community. Finally, we have given a series
of examples to demonstrate how the general visualization techniques of
focus-plus-context and linked brushing can be practically incorporated into a
range of practical analysis workflows, from studying the impact of bacteria on
human health to better allocating units in commuter bikesharing
systems.
